---
title: "Practice with t-tests"
output: html_document
date: "2022-10-27"
---

## Reviewing the t-test

Let's conduct a 2-sample t-test to compare body fat percentage between men and women.

The data set presents the body fat percentage for a sample of men and women who all work out at the same gym. In general, men are leaner than women, with a body fat percentage centering around 15% whereas for women it tends to center around 20%. Let’s see if the same is true of these data.

## Instructions

Use the information you learned from reading chapter 5 and doing homework 21 to conduct a t-test comparing mean body fat percentage between the male and female groups. But don’t just rush to the test – follow the recommended analysis workflow:

```{r}
rm(list = ls())
library(tidyverse)
library(here)

fat <- read.csv(here("data", "body_fat_percentage.csv"))
glimpse(fat)

fat$gender <- as.factor(fat$gender)
```


1.  Plot the data and look at expected relationships
```{r}
ggplot(fat, aes(x = body_fat_percent))+
  geom_histogram()+
  facet_wrap(~gender, ncol = 1)+
  theme_bw()
```

2.  Calculate the mean and standard deviation for the data and examine the resulting output.
```{r}
fat_stats <- fat |>
  group_by(gender) |>
  summarise(mean_body_fat = mean(body_fat_percent, na.rm = T),
            sd_body_fat = sd(body_fat_percent, na.rm = T))
```

3.  Discuss with your partner:
    -   Review in the book: What are the assumptions about the data appropriate for a t-test? When you examine the plots and summary data, do they appear to meet those assumptions?
    The data for male participants doesn't seem to be normally distributed.
    
    -   Based on examination of the plots and summary data, do you expect that there is a statistically significant difference in mean body fat percentage between males and females in this sample?
    Yes
    
4.  Conduct a t-test and then interpret the results. What do they tell you about the null vs. alternative hypotheses?
```{r}
t.test(body_fat_percent ~ gender, data = fat)
```

Men have significantly less body fat than women (t= , df= , p= ).