---
title: "21 Oct 2024 - review"
format: html
editor: visual
---

Download the dataset called "birds_and_mammals.csv" from today's module and put it in your Data folder.

Set up your workspace and read in the .csv. Name it "critters"

```{r}
rm(list = ls())
library(tidyverse)
library(here)

critters <- read.csv(here("data", "birds_and_mammals.csv"))
```

Filter just for mammals in the order Carnivora and make a new dataset called "carnivores"

```{r}
carnivores <- critters |>
  filter(Order == "Carnivora") 
  
```

create a histogram of maximum longevity for carnivores.

```{r}
ggplot(carnivores, aes(x = max_longevity_years))+
  geom_histogram()
```

Now make a scatterplot comparing max_longevity_years vs. adult_wt_g

```{r}
ggplot(carnivores, aes(x = max_longevity_years, y = adult_wt_g))+
  geom_point()
```

Use code (rather than scrolling the data) to determine which species it is that has the highest maximum longevity?

```{r}
carnivores |> 
  select(Common_name, max_longevity_years) |>
  arrange(desc(max_longevity_years))
```

Use code to determine the two species between 2.5e+05 and 3e+05 in adult body mass.

```{r}
carnivores |>
  select(Common_name, adult_wt_g) |>
  filter(adult_wt_g > 2.5e+05 &
           adult_wt_g < 3e+05)
```

Now make a boxplot of max longevity as a function of family. Set the y-axis label to "maximum longevity, years". Make the points red and somewhat transparent.

```{r}
ggplot(carnivores, aes(x = Family, y = max_longevity_years))+
  geom_boxplot()+
  geom_point(color = "red", alpha = 0.5)+
  labs( y = "maximum longevity, years")
```
